Reactive SINDy: Discovering governing reactions from concentration data

### 2018-12-05 docs/manuscript_jcp.tex
add figures with captions, add description to reaction tables

\subsection{Application to MAPK cascade\label{sec:mapk}}

\subsection{Application to Lotka-Volterra system\label{sec:lv}}

@@

\begin{table*}[h!]
    \centering %
    %\resizebox{.9\textwidth}{!}{%
    \begin{tabular}{rclcc}
        \toprule 
        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
        \midrule 
        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
        \bottomrule
    \end{tabular}\caption{\label{tab:reaction-library-mapk}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:mapk} for the MAPK system. The given rate constants define the ground truth reaction model.}
\end{table*}

\begin{table*}[h!]
    \centering %
    %\resizebox{.9\textwidth}{!}{%
    \begin{tabular}{rclcc}
        \toprule 
        \multicolumn{3}{c}{Reaction} & rate  & description \tabularnewline
        \midrule 
        $\mathrm{DNA}_{\mathrm{A}}$  & $\rightharpoonup$  & $\mathrm{DNA}_{\mathrm{A}}+\mathrm{mRNA}_{\mathrm{A}}$  & $k_{1}=1.8$  & transcription of $\mathrm{mRNA}_{\mathrm{A}}$\tabularnewline
        $\mathrm{mRNA}_{\mathrm{A}}+\mathrm{A}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{A}}$  & $k_{19}=0$  & artificial fusion\tabularnewline
        $\mathrm{mRNA}_{\mathrm{B}}+\mathrm{B}$  & $\rightharpoonup$  & $\mathrm{mRNA}_{\mathrm{B}}$  & $k_{20}=0$  & artificial fusion\tabularnewline
        \bottomrule
    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka-Volterra system. The given rate constants define the ground truth reaction model.}
\end{table*}

>>>>

\subsection{Application to MAPK cascade\label{sec:mapk}}
\begin{figure}
    \centering
    \includegraphics[width=\singlecolumnwidth]{figures_jcp/6.pdf}
    \caption{\label{fig:mapk}Application of reactive SINDy to the MAPK pathway system. \textbf{(a)}~The response curve of the MAPK cascade as a function of external stimulus given as a constant concentration $[\mathrm{S}]$. The activity is the long-time concentration of activated transcription factors $[\mathrm{TF*}]$. Dashed lines show the values of $[\mathrm{S}]$ at which concentration time series data was generated. \textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data. LSQ estimation and the ground truth model for comparison.}
\end{figure}

\subsection{Application to Lotka-Volterra system\label{sec:lv}}
\begin{figure}
    \centering
    \includegraphics[width=\singlecolumnwidth]{figures_jcp/7.pdf}
    \caption{\label{fig:lv}Application of reactive SINDy to the Lotka-Volterra system with social friction. \textbf{(a)}~Concentration data as a function of time for predator and prey. \textbf{(b)}~Estimated rate coefficients of candidate reactions after application of reactive SINDy (regularized) to the time series data. LSQ estimation and the ground truth model for comparison.}
\end{figure}

@@

        $\mathrm{S}        + \mathrm{MAPKKK}$  & $\rightharpoonup$  &  $\mathrm{S      } + \mathrm{MAPKKK*}$ & $k_{1} =1$ & external stimulus activates MAPKKK\tabularnewline
        $\mathrm{MAPKKK*}    \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK }                   $ & $k_{2} =1$ & dephosphorylation\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK* }$ & $k_{3} =1$ & phosphorylation of MAPKK\tabularnewline
        $\mathrm{MAPKK*}     \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  }                   $ & $k_{4} =1$ & dephosphorylation\tabularnewline
        $\mathrm{MAPKK*}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK*  }$ & $k_{5} =1$ & phosphorylation of MAPK\tabularnewline
        $\mathrm{MAPK*}      \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{MAPK   }                   $ & $k_{6} =1$ & dephosphorylation\tabularnewline
        $\mathrm{MAPK*}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF*    }$ & $k_{7} =1$ & phosphorylation of transcription factor\tabularnewline
        $\mathrm{TF*}        \mathrm{      }$  & $\rightharpoonup$  &  $\mathrm{TF     }                   $ & $k_{8} =1$ & dephosphorylation\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{MAPKK }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK* }$ & $k_{9} =0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  }                   $ & $k_{10}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF*    }$ & $k_{11}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK*  }$ & $k_{12}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF*    }$ & $k_{13}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK}    + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF*    }$ & $k_{14}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK*}   + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF*    }$ & $k_{15}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPK}     + \mathrm{TF    }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF*    }$ & $k_{16}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK}    + \mathrm{MAPK  }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK*  }$ & $k_{17}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPKK  }$ & $k_{18}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{MAPK   }$ & $k_{19}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK } + \mathrm{TF     }$ & $k_{20}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{MAPKK*}$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPKK  }$ & $k_{21}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{MAPK   }$ & $k_{22}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKKK*}  + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKKK*} + \mathrm{TF     }$ & $k_{23}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK}    + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{MAPK   }$ & $k_{24}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK  } + \mathrm{TF     }$ & $k_{25}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK*}   + \mathrm{MAPK* }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{MAPK   }$ & $k_{26}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPKK*}   + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPKK* } + \mathrm{TF     }$ & $k_{27}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPK}     + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK   } + \mathrm{TF     }$ & $k_{28}=0$ & artificial reaction\tabularnewline
        $\mathrm{MAPK*}    + \mathrm{TF*   }$  & $\rightharpoonup$  &  $\mathrm{MAPK*  } + \mathrm{TF     }$ & $k_{29}=0$ & artificial reaction\tabularnewline
    \end{tabular}\caption{\label{tab:reaction-library-mapk}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:mapk} for the MAPK system. The given rate constants define the ground truth reaction model.}

\begin{table*}[h!]
    \centering %
    %\resizebox{.9\textwidth}{!}{%
    \begin{tabular}{rclcc}
        \toprule 
        \multicolumn{3}{c}{Reaction} & rate & description \tabularnewline
        \midrule 
        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{1} =0.1$ & social friction of prey\tabularnewline
        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\emptyset             $  & $k_{2} =0.1$ & social friction of predator\tabularnewline
        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{3} =1  $ & prey growth\tabularnewline
        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{4} =1  $ & predator eats prey\tabularnewline
        $\mathrm{Y}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{5} =1  $ & predator decays\tabularnewline
        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X} + \mathrm{X}$ & $k_{6} =0  $ & artificial reaction\tabularnewline
        $\mathrm{X}             $ & $\rightharpoonup$ & $\emptyset             $  & $k_{7} =0  $ & artificial reaction\tabularnewline
        $\mathrm{Y} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{8} =0  $ & artificial reaction\tabularnewline
        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{9} =0  $ & artificial reaction\tabularnewline
        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{10}=0  $ & artificial reaction\tabularnewline
        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{11}=0  $ & artificial reaction\tabularnewline
        $\mathrm{X} + \mathrm{Y}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{12}=0  $ & artificial reaction\tabularnewline
        $\mathrm{X} + \mathrm{X}$ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{13}=0  $ & artificial reaction\tabularnewline
        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y}             $ & $k_{14}=0  $ & artificial reaction\tabularnewline
        $\mathrm{Y}             $ & $\rightharpoonup$ & $\mathrm{X}             $ & $k_{15}=0  $ & artificial reaction\tabularnewline
        $\mathrm{X}             $ & $\rightharpoonup$ & $\mathrm{Y} + \mathrm{Y}$ & $k_{16}=0  $ & artificial reaction\tabularnewline
        \bottomrule
    \end{tabular}\caption{\label{tab:reaction-library-lv}Full set of ansatz reactions $\Theta$ used in Sec.~\ref{sec:lv} for the Lotka-Volterra system. The given rate constants define the ground truth reaction model.}
\end{table*}
